---
title: "simpleRegress"
author: "mariahlynn woolfolk"
date: "summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

For model length~neck, adjusted $r^2$ = 0.75


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

## length based on age 

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

Adjusted $r^2$ for the age model = 0.51, this is less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.

## length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

Adjusted $r^2$ for the chest model = 0.79, this is greater than the adjusted $r^2$ for neck, so this model is better then the neck model.

## length based on sex 

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

Adjusted $r^2$ for the sex model = -0.01, this is less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.

## length based on head length

```{r}
headlengthModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlengthModel)
summary.lm(headlengthModel)
```

Adjusted $r^2$ for the head length model = 0.84, this is greater than the adjusted $r^2$ for neck, so this model is better then the neck model.

## length based on head width

```{r}
headwidthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwidthModel)
summary.lm(headwidthModel)
```

Adjusted $r^2$ for the head width model = 0.53, this is less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.

## length based on weight

```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```

Adjusted $r^2$ for the weight model = 0.74, this is less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.

### Conclusion

Overall, the best model is LENGTH~HEADLEN because it has the greatest adjusted $r^2$ of 0.84, while the other models had much lower adjusted $r^2$'s








